### 2021-04-25 Paper/Main.tex

\section{Literature Review}
This section will be broke into two components, we will first look at theoretical models for asset purchasing and then we will introduce some of the mythology for introducing a EGS factor into the asset pricing framework.
to account for the ESG factor of a five factor famma french model, the authers used the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index creating a facotr by subtracting the retunrs from the high ESG portfolio from the low ESG portfolio. \cite{Gregory2020TheGP}
\label{Literature}

@@

An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index. Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors. The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis. In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}. Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium. One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio. However \$Vice has only traded for two years, as such the data is far less robust. Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.

@@

For this analysis, daily returns data was collected between June 2011 and March 2021. The started was determined by the initiation of the MSCI ESG index. To collect daily individual stock return information CRSP data parsed from a larger dataset and then collapsed on month by summing each individual days returns to creat monthly returns. Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures. Finally the EMP factor was also  constructed summing the daily returns for each much. All three of the datasets were merged using the date and month as the matching component. The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security. Below in table 1 are the summary statistics for the raw data.

@@

To capture abnormal returns we will consider 5 factors that is the 5 factors from the famma french paper small minus big(SMB),CMA, HML, and market portfolio(I.e model, consisting of market, size, value, profitability, and investment factors;). In addition I will ad another term to address how stocks move with the movement to a ESG porlfio. For this purpose we will use the MSCI ESG all world index. 
\section{Robustness Checks}
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
\label{Robustness}

@@

\vspace{1cm}
\textbf{Table 2:} This table shows many things\\
\vspace{1cm}

>>>>

\section{Literature Review}
\label{Literature}
This section will be broke into two components. The first section will adress  theoretical models for asset purchasing and then we will introduce some of the mythology for  EGS factor in the asset pricing framework.\\

 \hl{add a section on the metholoogy for the 2 pass sort use} \cite{Kleibergen2013UnexplainedFA} and \cite{Fama1973RiskRA} \\
The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964's work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}. The CAPM model was later iterated on by Fama french in 1993 with the 3 factor, which considered several other terms to predict stock returns. Fama and French added size and book to market terms to the CAPM model. They found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies(low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}. Fama and French again innovated on the three-factor model in 2015 by adding two factors to the regression. In this model, Fama and French consider Investment and profitability while holding the other regression stable factors. The five-factor model analysis found that the addition of the investment and profitability factor made the value factor redundant \cite{Fama2013AFA}. The five-factor framework analysis has been recreated on international exchanges to test the effects on foreign markets. In recreation by Fama French and another study by Zhong and Li, the general relationships demonstrated in the five-factor model are also applicable in China, Japan, and Australian markets \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}. However, when the Fama French regression was applied to the Australian equity market, the authors did not find the significance of the value factor to subside like was visible in the 2015 paper \cite{Chiah2016A}.  In addition to expanding the five-factor analysis geographically, some scholars have expanded the five-factor model by adding additional terms to assess momentum effects.  In a paper published by  Dirkx and  Peter, the authors expanded the five-factor model by applying the Fama French framework to the German market while also adding a sixth term to assess the effect of risk premium with the CDAX constituents' momentum. Another study by Gregory and Stead used the sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia. This paper found that  "the risk premium of the sustainability factor is positive and significant\cite{Gregory2020TheGP}."

In addition to Gregory and Stead's paper, several other authors have used the Fama French model to address the question concerning ESG and risk premia. To set up an ESG Fama French model, many academic journal articles use an index as a positive ESG portfolio and another index as a negative portfolio to create the ESG factor. In Gregory and Stead's paper, the author uses the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index to create a factor by subtracting the returns from the high ESG portfolio from the low ESG portfolio \cite{Gregory2020TheGP}. In Dorfleitner et al.'s paper, the authors collect Thomas Reuter's controversies and combined ESG score for 2500 companies. With the ranked ESG score data, the author then creates two portfolios by separating the top decile score and the bottom deciles scores, which act as the ESG and contra ESG portfolios. Similar to Gregory et al. paper, the author subtract the top 10\% from the bottom 10\% to create the ESG factor \cite{Dorfleitner2020ESGCA}.

@@

An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index. Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors. The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis. In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}. Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium. One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio. However \$Vice has only traded for two years, as such the data is far less robust. Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.

@@

For this analysis, daily returns data was collected between June 2011 and March 2021. The started was determined by the initiation of the MSCI ESG index. To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns. Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures. Finally the EMP factor was also  constructed summing the daily returns for each much. All three of the datasets were merged using the date and month as the matching component. The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security. Below in table 1 are the summary statistics for the raw data.

@@

\section{Robustness Checks}
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
\label{Robustness}

@@

\section{Appendix}
\vspace{1cm}
\textbf{Table 2:} This table shows the mean results by decile before any regressions\\
\vspace{1cm}
